Do return series have power-law tails with the same index?

### 2017-04-12 papers/FX/draft.tex
Revised the FX manuscript.

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
%\usepackage[bookmarks=true]{hyperref}
%\usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

$\E [(\alpha_1 Z_1+\beta_1)^{\alpha/2}]=1$

@@

\subsection{An utility based argument for equal tail indices} \setcounter{equation}{0}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red meaning of these constants?} 
 Here $v$ can be thought of as the certainty payoff equivalent to the
risky payoff $C$. If $b=0$, 
preferences are the classical expected utility. If $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin \cite{routledge2010generalized}.
An agent guided by the utility function $u$ will seek to maximize the \fct al $\wt u$.
Routledge and Zin assumed
a power-law utility function 
\beam\label{eq:u}
  u(x)=\frac{1}{-\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
\eeam
For the sake of argument, we assume that an investor initially has one unit of
wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with return $X$ over one time unit, i.e.,
\beao%\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeao
Then we have 
\beao
\wt u(\phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C < \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \ln \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C < \delta v$ \fif\ $X < q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is a functional of $\phi$ and $F_X$,
$
\wt u=\wt u (F_X, \phi)\,,
$
and we assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
{\red Is this worth a theorem?}
\begin{theorem}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
\end{theorem}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of Theorem~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}. {\red how can I see positive dependence?}
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$
    (left) and $\xi = 4$ (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other figures, and use the same
structure of the formulas.}
  }
  \label{fig:phi_hat_pareto}
\end{figure}

@@

\subsection{Hybrid fitting of the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:X_distr}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:X_distr}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$. Regularity
conditions on the maximum likelihood estimator are easily checked for
the distribution \eqref{eq:X_distr}.
The density function of $X$ conditional on $X < 0$ is
\[
f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
\]
Clearly, the log-likelihood function of the $K$ parameter is
\[
l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
\]
where, as is just mentioned, we pre-estimate $\alpha$ with Hill
estimator and regard it as known at this step. With $n$ observations
$X_1, \dots, X_n$, the log-likelihood function of $K$ is
\[
L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
\]
To obtain the maximum likelihood estimation, we need to solve the
equation ${\pd L \over \pd K} = 0$, which expands to
\[
{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
\]
This equation has to be solved numerically.
The Fisher information of $K$ follows as
\begin{eqnarray*}
  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
  &=& \alpha K^{\alpha - 2}
  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
  &=& {\alpha \over (2 + \alpha) K^2}
\end{eqnarray*}
So the maximum likelihood estimator, call it $\hat K$, is
asymptotically normal:
\[
\sqrt n (\hat K - K) \to N \left(
  0, \left( 1 + {2 \over \alpha} \right) K^2
\right)
\]

To obtain an estimate of $p$, we use the law of large number:
\begin{eqnarray*}
  p &=& \P(X < 0) = \lim_{n \to \infty}
  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
\end{eqnarray*}
By the central limit theorem,
\begin{eqnarray*}
  \sqrt n \left[
    {1 \over n}\sum_{t=1}^n \1{X_t < 0} - p
  \right] \to N(0, p - p^2))
\end{eqnarray*}

@@

\subsubsection{log-utility function}
We first consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:power_utility} as $\gamma \to 0$.

@@

\appendix
\section{Proof of lemma \ref{lemma:I}}
\begin{proof}
  Clearly
  \begin{eqnarray*}
    && {\pd \E h(X) \over \pd \theta} \\
    &=& \int_a^b h(x) {\pd f \over \pd \theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_2}
  \end{eqnarray*}
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd \theta}(x,
    \theta) > 0$ on $(a, x_0)$
    \[
    I_1 > h(x_0) \int_a^{x_0} {\pd f \over \pd \theta}(x, \theta) dx
    \]
    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
    $h(x)$ is decreasing, we have
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx \\
      &>& -h(x_0)
      \int_{x_0}^b \left| 
        {\pd f \over \pd \theta}(x, \theta)
      \right| dx
    \end{eqnarray*}
    Finally we have
    \begin{eqnarray*}
      {\pd \E h(X) \over \pd \theta}
      > h(x_0) \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
      = h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx
      = 0
    \end{eqnarray*}
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    ${\pd f \over \pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    ${\pd f \over \pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \begin{eqnarray*}
      I_1 &=&
      \int_a^{x_0} -h(x)
      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx \\
      &>&
      -h(x_0) \int_a^{x_0}
      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
    \end{eqnarray*}
    and obviously
    \[
    I_2 > h(x_0) \int_{x_0}^b
    {\pd f \over \pd \theta}(x, \theta) dx
    \]
    So
    \[
    {\pd \E h(X) \over \pd \theta}
    = I_1 + I_2
    > h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx = 0
    \]
  \end{enumerate}
\end{proof}

>>>>

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
%% \usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$

@@

\subsection{An utility based argument for equal tail indices} \setcounter{equation}{0}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
  meaning of these constants?}
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(\phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C < \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \ln \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C < \delta v$ \fif\ $X < q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is a functional of $\phi$ and $F_X$,
$
\wt u=\wt u (F_X, \phi)\,,
$
and we assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
    {\red Is this worth a theorem?}
    {\blue No}
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
\end{remark}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}. {\red how can I see positive dependence?}
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$
    (left) and $\xi = 4$ (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other figures, and use the same
structure of the formulas.}
  }
  \label{fig:phi_hat_pareto}
\end{figure}

@@

\subsection{Fitting the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:pareto}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:pareto}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$. Regularity
conditions on the maximum likelihood estimator are easily checked for
the distribution \eqref{eq:pareto}.
The density function of $X$ conditional on $X < 0$ is
\[
f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
\]
Clearly, the log-likelihood function of the $K$ parameter is
\[
l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
\]
where, as is just mentioned, we pre-estimate $\alpha$ with Hill
estimator and regard it as known at this step. With $n$ observations
$X_1, \dots, X_n$, the log-likelihood function of $K$ is
\[
L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
\]
To obtain the maximum likelihood estimation, we need to solve the
equation ${\pd L \over \pd K} = 0$, which expands to
\[
{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
\]
This equation has to be solved numerically.
The Fisher information of $K$ follows as
\begin{eqnarray*}
  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
  &=& \alpha K^{\alpha - 2}
  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
  &=& {\alpha \over (2 + \alpha) K^2}
\end{eqnarray*}
So the maximum likelihood estimator, call it $\hat K$, is
asymptotically normal:
\[
\sqrt n (\hat K - K) \to N \left(
  0, \left( 1 + {2 \over \alpha} \right) K^2
\right)
\]

To obtain an estimate of $p$, we use the law of large number:
\begin{eqnarray*}
  p &=& \P(X < 0) = \lim_{n \to \infty}
  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
\end{eqnarray*}
By the central limit theorem,
\begin{eqnarray*}
  \sqrt n \left[
    {1 \over n}\sum_{t=1}^n \1_{\{X_t < 0\}} - p
  \right] \to N(0, p - p^2))
\end{eqnarray*}

@@

\subsubsection{log-utility function}
We first consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:u} as $\gamma \to 0$.

@@

\appendix
\section{A monontonicity lemma}
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
  $\E h(X) < \infty$ and
  \begin{eqnarray*}
    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
    \quad \forall x \in (a, b)
  \end{eqnarray*}
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \end{enumerate}
\end{lemma}
\begin{remark}
  \label{remark:I}
  Two other cases follow trivially from lemma \ref{lemma:I}:
  \begin{enumerate}
  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
    satisfies the
    same conditions of the 1st case of lemma \ref{lemma:I},
    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
  \item By the same argument, if $h(\cdot)$ is decreasing and
    ${\pd f \over \pd \theta}$ satisfies the same conditions of the
    2nd case of lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
  \end{enumerate}
\end{remark}

\begin{proof}
  Clearly
  \begin{eqnarray*}
    && {\pd \E h(X) \over \pd \theta} \\
    &=& \int_a^b h(x) {\pd f \over \pd \theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x,
      \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x,
      \theta) dx}_{I_2}
  \end{eqnarray*}
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd
    \theta}(x, \theta) > 0$ on $(a, x_0)$
    \[
    I_1 > h(x_0) \int_a^{x_0} {\pd f \over \pd \theta}(x, \theta) dx
    \]
    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for
    $x \in (x_0, b)$ and $h(x)$ is decreasing, we have
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx \\
      &>& -h(x_0)
      \int_{x_0}^b \left| 
        {\pd f \over \pd \theta}(x, \theta)
      \right| dx
    \end{eqnarray*}
    Finally we have
    \begin{eqnarray*}
      {\pd \E h(X) \over \pd \theta}
      > h(x_0) \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
      = h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx
      = 0
    \end{eqnarray*}
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    ${\pd f \over \pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    ${\pd f \over \pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$
    \begin{eqnarray*}
      I_1 &=&
      \int_a^{x_0} -h(x)
      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx \\
      &>&
      -h(x_0) \int_a^{x_0}
      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
    \end{eqnarray*}
    and obviously
    \[
    I_2 > h(x_0) \int_{x_0}^b
    {\pd f \over \pd \theta}(x, \theta) dx
    \]
    So
    \[
    {\pd \E h(X) \over \pd \theta}
    = I_1 + I_2
    > h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx = 0
    \]
  \end{enumerate}
\end{proof}
